---
title: "CRIMES IN BOSTON ANALYSIS"
author: "Arya Andhika"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
#Library Needed
library(lubridate)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggmosaic)
library(scales)
library(RColorBrewer)

#Reading Data
boston <- read.csv("Data Source/crime.csv")
master.boston <- boston
#boston <-  master.boston #when need to recall data

#Set some colors
cbPalette <- c("steelblue2", "springgreen2", "yellow2", "brown2")
```

## Take A Look the Data Type of "Crime in Boston" Data"
```{r}
str(boston)
```

## Change the Data Type of "OCCURRED_ON_DATE" 
```{r}
boston$INCIDENT_NUMBER <- as.character(boston$INCIDENT_NUMBER) #change the type into chr because unique
boston$OCCURRED_ON_DATE <- ymd_hms(boston$OCCURRED_ON_DATE,tz = "America/New_York") # change into date format
```

## Change the Data Type of "OFFENSE_CODE" and "REPORTING_AREA" into Factor (Repeated)
```{r}
boston$OFFENSE_CODE <- as.factor(boston$OFFENSE_CODE)
boston$REPORTING_AREA <- as.factor(boston$REPORTING_AREA)
```

## Take A Look the Summary of "Crime in Boston" Data"
```{r}
summary(boston)
```

## Clean Up the "NA's" and "Blank" from Data
```{r}
boston <- boston[complete.cases(boston), ]
boston <- boston[!(boston$DISTRICT == ""), ]
```

## Change Month form numeric format into Abb. Month and Reorder Them
```{r}
boston$MONTH <- as.factor(month.abb[boston$MONTH])
boston$MONTH <- factor(boston$MONTH, levels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) #reordering Level Month
```

## DAY_OF_WEEK stored in long format and not orederd

```{r}
boston$DAY_OF_WEEK <- wday(boston$OCCURRED_ON_DATE,
                           label = T,
                           abbr = T,
                           week_start = 1)
```

## Replace Missing Values in "SHOOTING" with "N"
```{r}
boston$SHOOTING <- as.factor(sub("^$", "N", boston$SHOOTING))
```

## Clean Up the Long and Lat Anomalies 
```{r}
boston <- boston[!(boston$Lat == -1 & boston$Long == -1), ]
```

## Remove unused levels
```{r}
boston <- droplevels(boston) 
```


##Simple Exploratory Data Analysis

Take a glimpse look at Crime Boston Data to find pattern or interesting findings


### By Year
```{r}
ggplot(boston, aes(x = YEAR)) +
  geom_bar(fill = "royalblue2", col = "mediumblue") +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston",
       subtitle = "During Year of 2015 - 2019") +
  geom_text(aes(label=comma(..count..)),stat="count", position=position_dodge(0.9),vjust=-0.5) +
  scale_y_continuous(labels=comma)
```

We can see from the chart that Crimes in Boston Occured mostly in 2016 - 2017


### By Month
```{r}
ggplot(boston, aes(x = MONTH)) +
  geom_bar(fill = "royalblue2", col = "mediumblue") +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston by Month",
       subtitle = "During Year of 2015 - 2019") +
   geom_text(aes(label=comma(..count..)),stat="count", position=position_dodge(0.9),vjust=-0.5) +
   scale_y_continuous(labels=comma)
```

Crimes in Bostod peaked at July


### By Season

Since there are 4 seasons in US, it will be interesting to see whether the seasons may affect the Crimes Rate
```{r}

# Create Function to divide 12 months into 4 seasons
season <- function(m){
                      if(m == "Mar" | m == "Apr" | m == "May"){
                        m <- "SPRING"
                      }else if(m == "Jun" | m == "Jul" | m == "Aug"){
                        m <- "SUMMER"
                      }else if(m == "Sep" | m == "Oct" | m == "Nov"){
                        m <- "AUTUMN"
                      }else{
                        m <- "WINTER"
                      }
                      
                    }

# Apply it to Month and create new Column
boston$SEASON_OCCURED <- as.factor(sapply(boston$MONTH, season))

# Reorder it
boston$SEASON_OCCURED <- ordered(boston$SEASON_OCCURED,
                                                   levels = c("WINTER",
                                                              "SPRING",
                                                              "SUMMER",
                                                              "AUTUMN"))

ggplot(boston, aes(x = MONTH, fill = SEASON_OCCURED)) +
  geom_bar() +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston by Season",
       subtitle = "During Year of 2015 - 2019",
       fill = "Season") +
   geom_text(aes(label=comma(..count..)),stat="count", position=position_dodge(0.9), vjust=-0.5, size = 3) +
   scale_y_continuous(labels=comma) +
   scale_fill_manual(values = cbPalette[1:4]) +
   theme(legend.position = "bottom",
         legend.title = element_text(size = 8),
         legend.text = element_text(size = 8))

```

Summer is the season where Crimes in Boston Mostly Occured

### By Day of Week
```{r}
ggplot(boston, aes(x = DAY_OF_WEEK)) +
  geom_bar(fill = "royalblue2", col = "mediumblue") +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston by Day of Week",
       subtitle = "During Year of 2015 - 2019") +
   geom_text(aes(label=comma(..count..)),stat="count", position=position_dodge(0.9),vjust=-0.5) +
   scale_y_continuous(labels=comma)
```

Crimes in Boston occur mostly in weekday, and peaked during Friday

### By Time
```{r}
event <- function(x){
                      if(x < 6){
                        x <- "12AM to 6AM"
                      }else if(x >= 6 & x < 12){
                        x <- "6AM to 12PM"
                      }else if(x >= 12 & x < 18){
                        x <- "12PM to 6PM"
                      }else{
                        x <- "6PM to 12AM"
                      }
                      
                    }

boston$TIME_OCCURED <- as.factor(sapply(boston$HOUR, event))

boston$TIME_OCCURED <- ordered(boston$TIME_OCCURED,
                                                   levels = c("12AM to 6AM",
                                                              "6AM to 12PM",
                                                              "12PM to 6PM",
                                                              "6PM to 12AM"))

ggplot(boston, aes(x = TIME_OCCURED)) +
  geom_bar(fill = "royalblue2", col = "mediumblue") +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston by Time Occured",
       subtitle = "During Year of 2015 - 2019") +
   geom_text(aes(label=comma(..count..)),stat="count", position=position_dodge(0.9),vjust=-0.5) +
   scale_y_continuous(labels=comma)
```

Crimes happen mostly at 12PM - 6PM


### Correlation between Day of Week and the Time when Crimes Occur
```{r}
ggplot(boston) +
  geom_mosaic(aes(x = product(DAY_OF_WEEK), fill=TIME_OCCURED)) +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston by Day of Week vs Time Occured",
       subtitle = "During Year of 2015 - 2019",
       fill = "Time Occured") +
  theme(legend.position = "bottom",
         legend.title = element_text(size = 8),
         legend.text = element_text(size = 8))
```

On the weekday, Crimes in Boston happen during the same period. But interesting thing happen at weekend, there is some shifting of time occured, from 6AM-12PM into 12AM to 6AM.


## Take a look the most occured Offense Type, let's say they occur more than 10k Cases
```{r}
boston10k <- as.data.frame(table(OFFENSE_TYPE = boston$OFFENSE_CODE_GROUP))
boston10k <- boston10k[boston10k$Freq >= 10000, ]
boston10k <- boston10k[order(boston10k$Freq, decreasing = T),]
boston10k
```

```{r}
boston10k <- boston10k[boston10k$OFFENSE_TYPE != "Other", ]
```

We would like to take out Other since it may consists of few Offense Types


## Plot It Into Chart
```{r}
ggplot(boston10k, aes(x = reorder(OFFENSE_TYPE, Freq), y = Freq)) +
  geom_col(aes(fill=OFFENSE_TYPE)) +
  coord_flip() +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Top 10 Offense Type Numbers in Boston Crimes",
       subtitle = "During Year of 2015 - 2019") +
  scale_fill_brewer(palette = "Paired") +
  theme(legend.position = "none") +
  scale_y_continuous(labels=comma)
```


## Location Analysis

### By Districts
```{r}

newlev <- names(table(boston$DISTRICT))[order(table(boston$DISTRICT), decreasing = T)]
boston$DISTRICT <- factor(boston$DISTRICT, levels=newlev)

ggplot(boston, aes(x = DISTRICT)) +
  geom_bar(fill = "royalblue2", col = "mediumblue") +
  theme_igray() +
  labs(x = NULL, y = NULL,
       title = "Number of Crimes in Boston on Every Districts",
       subtitle = "During Year of 2015 - 2019") +
   geom_text(aes(label=comma(..count..)),stat="count", position=position_dodge(0.9),vjust=-0.5) +
   scale_y_continuous(labels=comma)
```

B2 is the District with most Cases, and A15 is the safest


### Districs vs Offense Type
We would like to see, the distribution of Top Numbers Offense Type (more than 10k cases) in every districts

```{r}
bostonsort <- boston[which(boston$OFFENSE_CODE_GROUP %in% boston10k$OFFENSE_TYPE), ] # subset boston dataset which only contains Top Offense Type
bostonsort <- droplevels(bostonsort) # remove unused levels

sortlev <- names(table(bostonsort$DISTRICT))[order(table(bostonsort$DISTRICT), decreasing = T)] #sort form largest - smallest
bostonsort$DISTRICT <- factor(bostonsort$DISTRICT, levels=sortlev) # set as factor

ggplot(bostonsort) +
  geom_mosaic(aes(x = product(DISTRICT), fill=OFFENSE_CODE_GROUP)) +
  labs(x = NULL, y = NULL,
       title = "Crimes in Boston by Offense Type in Every Districts",
       subtitle = "During Year of 2015 - 2019",
       fill = "Time Occured") +
  theme_igray() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 7)) +
  scale_fill_brewer(palette = "Paired")


```

- As we can see, the order of the District is the same with the unsorted one. 
- Motor Vehicle Accident Response is the most occured case on every distrcit
- Larceny should be main concern at D4 and A1


### Number of Crimes on Every Date

```{r}
#Change Format Date from POSIXct into Date
bostonsort$OCCURRED_ON_DATE <- as.Date(as.POSIXct(bostonsort$OCCURRED_ON_DATE,tz = "America/New_York"))


#count numbers of crimes by date of occurance
bostoncount <- bostonsort %>% 
                  group_by(OCCURRED_ON_DATE) %>% 
                  summarise(count = n()) %>% 
                  ungroup()

ggplot(bostoncount, aes(x = OCCURRED_ON_DATE, y = count)) +
  geom_point(col = "royalblue2") +
  labs(x = NULL, y = NULL,
       title = "Count of Crimes in Boston by Date of Occurance",
       subtitle = "During Year of 2015 - 2019") +
  theme_igray()

```

Same like as shown by Month, the numbers of crimes mostly occured on mid year, and decline at the end of the year